---
title: "Data Summaries Demo"
author: "Dave Paunesku"
date: "September, 2016"
output: 
  html_document:
    fig_caption: no
    toc: yes
    theme: spacelab
    css: https://raw.githubusercontent.com/PERTS/gymnast/master/Rmd/rmd_styles.css
---


## Purpose

The data summaries (`ds`) module in gymnast is used to easily summarize data. This demo file shows how to use it.



## Load Gymnast 
```{r, results='hide', message=FALSE, warning=FALSE}

options(stringsAsFactors = FALSE)
github_base_path <- "https://raw.githubusercontent.com/PERTS/gymnast/master/"

tryCatch({
    source("~/Sites/gymnast/R/util.R")
    source("~/Sites/gymnast/R/util_data_summaries.R")
}, error = function(e){
    source(github_base_path %+% "R/util.R")
    source(github_base_path %+% "R/util_data_summaries.R")
})

```

## Load Sample Data
Load a sample data file `data` and an optional codebook file `codebook`.


```{r, results='hide', message=FALSE, warning=FALSE}

tryCatch({
    codebook <- read.csv("~/Sites/gymnast/data/skill-aquirability-variable_key.csv")
    data <- read.csv("~/Sites/gymnast/data/skill-aquirability.csv")
}, error = function(e){
    codebook <- 
        read.csv( github_base_path %+% "data/skill-aquirability-variable_key.csv")
    data <- 
        read.csv( github_base_path %+% "data/skill-aquirability")
})

```

## Display Data
To give a flavor of what the codebook and data look like.

```{r, message=FALSE, warning=FALSE, results='asis'}

data[,1:10] %>% 
    head() %>%
    util.html_table()

codebook %>% 
    head() %>%
    util.html_table()

```

## Using the ds.summarize_by_column function
The summarize by column (`ds.summarize_by_column`) function is used to create
univariate descriptions of data. 

### Default Behavior
By default, `ds.summarize_by_column` uses the functions in
`ds.helper$default_col_funcs` to summarize data.

```{r, message=FALSE, warning=FALSE, results='asis'}

ds.summarize_by_column(data) %>% 
    head() %>%
    util.html_table()

```

### Append a Codebook
If you pass a codebook in the `codebook` argument,
`ds.summarize_by_column` will automatically merge it in on 
the `variable_name` column. Columns from `data`
are retained even if they are not matched in the 
codebook.

```{r, message=FALSE, warning=FALSE, results='asis'}

codebook_columns_to_include <- c("variable_name","variable_label")
ds.summarize_by_column(data, codebook=codebook[, codebook_columns_to_include]) %>% 
    head() %>%
    util.html_table()

```

### Customize Summary Variables
Supply your own summary functions to `ds.summarize_by_column` via the 
`func_list` argument if you want to change the summary.
For example, here we include the mean, standard deviation,
and standard error.

```{r, message=FALSE, warning=FALSE, results='asis'}

custom_func_list <- list(
    "mean" = ds.helper$mean,
    "sd" = ds.helper$sd,
    "se" = function(x){
        # add in the standard error
        x <- x[!util.is_blank(x)]
        sd(x)/length(x)**.5
    }
)

ds.summarize_by_column(data, 
            func_list = custom_func_list,
            codebook=codebook[, codebook_columns_to_include]) %>% 
    head() %>%
    util.html_table()

```

### Use Default Categorical Descriptives
Default description functions for categorical variables are provided in 
`ds.helper$default_categorical_col_funcs` and can be passed to 
the `func_list` argument to summarize categorical data.

```{r, message=FALSE, warning=FALSE, results='asis'}

# select the categorical columns for description
categorical_columns <- ! util.apply_columns(data, util.is_vector_of_numbers)
ds.summarize_by_column( 
        data[ , categorical_columns ], 
        func_list=ds.helper$default_categorical_col_funcs) %>% 
    head() %>%
    util.html_table()

```






















