---
title: "Data Summaries Demo"
author: "Dave Paunesku"
date: "September, 2016"
output: 
  html_document:
    fig_caption: no
    toc: yes
    theme: spacelab
    css: https://raw.githubusercontent.com/PERTS/gymnast/master/Rmd/rmd_styles.css
---


## Purpose

The data summaries (`ds`) module in gymnast is used to easily summarize data. This demo file shows how to use it.



## Load Gymnast 
```{r, results='hide', message=FALSE, warning=FALSE}

options(stringsAsFactors = FALSE)
github_base_path <- "https://raw.githubusercontent.com/PERTS/gymnast/master/"

tryCatch({
    source("~/Sites/gymnast/R/util.R")
    source("~/Sites/gymnast/R/util_data_summaries.R")
}, error = function(e){
    source(github_base_path %+% "R/util.R")
    source(github_base_path %+% "R/util_data_summaries.R")
})

```

## Load Sample Data
Load a sample data file `data` and an optional codebook file `codebook`.


```{r, results='hide', message=FALSE, warning=FALSE}

tryCatch({
    codebook <- read.csv("~/Sites/gymnast/data/skill-aquirability-variable_key.csv")
    data <- read.csv("~/Sites/gymnast/data/skill-aquirability.csv")
}, error = function(e){
    codebook <- 
        read.csv( github_base_path %+% "data/skill-aquirability-variable_key.csv")
    data <- 
        read.csv( github_base_path %+% "data/skill-aquirability")
})

```

## Display Data
To give a flavor of what the codebook and data look like.

```{r, message=FALSE, warning=FALSE, results='asis'}

data[,1:10] %>% 
    head() %>%
    util.html_table()

codebook %>% 
    head() %>%
    util.html_table()

```

## Using the ds.summarize_by_column function
The summarize by column (`ds.summarize_by_column`) function is used to create
univariate descriptions of data. 

### Default Behavior
By default, `ds.summarize_by_column` uses the functions in
`ds.helper$default_col_funcs` to summarize data.

```{r, message=FALSE, warning=FALSE, results='asis'}

ds.summarize_by_column(data) %>% 
    head() %>%
    util.html_table()

```

### Append a Codebook
If you pass a codebook in the `codebook` argument,
`ds.summarize_by_column` will automatically merge it in on 
the `variable_name` column. Columns from `data`
are retained even if they are not matched in the 
codebook.

```{r, message=FALSE, warning=FALSE, results='asis'}

codebook_columns_to_include <- c("variable_name","variable_label")
ds.summarize_by_column(data, codebook=codebook[, codebook_columns_to_include]) %>% 
    head() %>%
    util.html_table()

```

### Customize Summary Variables
Supply your own summary functions to `ds.summarize_by_column` via the 
`func_list` argument if you want to change the summary.
For example, here we include the mean, standard deviation,
and standard error.

```{r, message=FALSE, warning=FALSE, results='asis'}

custom_func_list <- list(
    "mean" = ds.helper$mean,
    "sd" = ds.helper$sd,
    "se" = function(x){
        # add in the standard error
        x <- x[!util.is_blank(x)]
        sd(x)/length(x)**.5
    }
)

ds.summarize_by_column(data, 
            func_list = custom_func_list,
            codebook=codebook[, codebook_columns_to_include]) %>% 
    head() %>%
    util.html_table()

```

### Use Default Categorical Descriptives
Default description functions for categorical variables are provided in 
`ds.helper$default_categorical_col_funcs` and can be passed to 
the `func_list` argument to summarize categorical data.

```{r, message=FALSE, warning=FALSE, results='asis'}

# select the categorical columns for description
categorical_columns <- ! util.apply_columns(data, util.is_vector_of_numbers)
ds.summarize_by_column( 
        data[ , categorical_columns ], 
        func_list=ds.helper$default_categorical_col_funcs) %>% 
    head() %>%
    util.html_table()

```

## Building Models
The `ds` module creates and runs models based on desired
outcomes and predictors.

### Build Formulas with ds.build_formulas()
`ds.build_formulas()` builds a vector of model formulas for 
each unique combination of the `dv` x `iv` x `mod` x `cov_group`. 
If `covs` is supplied, cov_group gets one entry for no covs (unadjusted) 
and one for all covs supplied in covs argument (adjusted).

```{r, message=FALSE, warning=FALSE, results='asis'}
dvs = c("sksh_post","bls_link") # dependent variables each run independently
ivs = c("condition","skil_aq_1") # independent variables each run independently
mods = c("","in_job_satis")  # moderators each run independently
cov_groups = list(
    "unadjusted" = c(),
    "demographic_adjustment" = c("age","male","race"),
    "demog_and_sksh_pre" = c("sksh_pre","age","male","race")
)  # vectors of covariates to run together
# covs = c()         # covariates included all together or none
                     # here, we use cov_groups instead for more flexibility

ds.build_formulas( dvs=dvs, ivs=ivs, mods=mods, cov_groups = cov_groups) %>%
    util.print_pre()

```

### Run GLMs with ds.glms() then print table with ds.glm_table
This section models how to use `ds.glms()` and `ds.glms_table`.
Note that ds.glm_table only work with models that meet the
ds.glm1 spec: 

1. IVs and moderators must be boolean or numeric
2. No more than one interaction via "*", though you can provide manually


```{r, message=FALSE, warning=FALSE, results='asis'}

dvs = c("sksh_post","bls_link") # dependent variables each run independently
ivs = c("treatment") # independent variables each run independently
mods = c("","job_satis")  # moderators each run independently, default no mod
cov_groups = list(
    "unadjusted" = c(),
    "demographic_adjustment" = c("age","male","race")
)  # vectors of covariates to run together
# covs = c()         # covariates included all together or none
                     # here, we use cov_groups instead for more flexibility
data$treatment <- data$condition %in% "treatment"

formulas <- ds.build_formulas( 
                dvs=dvs, ivs=ivs, mods=mods, cov_groups = cov_groups
            )

models   <- ds.glms(formulas, data)

# create a more concise table with just the apa strings 
ds.glm_table(models, columns=ds.helper$glm1_features_apa) %>% 
    util.html_table()

# show the full table to see all the reported features
ds.glm_table(models, columns=ds.helper$glm1_features_full) %>% 
    util.html_table()

```

### Example of off-spec model
When off-spec models are requested, ds.glm_table issues warnings and 
omits them from the table.

```{r, message=FALSE, warning=FALSE, results='asis'}

dvs = c("sksh_post") # dependent variables each run independently
ivs = c("condition","treatment") # independent variables each run independently
mods = c("","job_satis")  # moderators each run independently, default no mod
cov_groups = list()  # vectors of covariates to run together
# covs = c()         # covariates included all together or none
                     # here, we use cov_groups instead for more flexibility
data$treatment <- data$condition %in% "treatment"

formulas <- ds.build_formulas( 
                dvs=dvs, ivs=ivs, mods=mods, cov_groups = cov_groups
            )

models   <- ds.glms(formulas, data)

# create a more concise table with just the apa strings 
ds.glm_table(models, columns=ds.helper$glm1_features_apa) %>%
    util.html_table()


```


















