---
title: "Data Summaries Demo"
author: "Dave Paunesku"
date: "September, 2016"
output: 
  html_document:
    fig_caption: no
    toc: yes
    theme: spacelab
    css: https://raw.githubusercontent.com/PERTS/gymnast/master/Rmd/rmd_styles.css
---


## Purpose

The data summaries (`ds`) module in gymnast is used to easily summarize data. This demo file shows how to use it.



## Load Gymnast 
```{r, results='hide', message=FALSE, warning=FALSE}

options(stringsAsFactors = FALSE)
github_base_path <- "https://raw.githubusercontent.com/PERTS/gymnast/master/"

tryCatch({
    source("~/Sites/gymnast/R/util.R")
    source("~/Sites/gymnast/R/util_data_summaries.R")
    source("~/Sites/gymnast/R/util_graphing.R")
}, error = function(e){
    source(github_base_path %+% "R/util.R")
    source(github_base_path %+% "R/util_data_summaries.R")
    source(github_base_path %+% "R/util_graphing.R")
})

```

## Load Sample Data
Load a sample data file `data` and an optional codebook file `codebook`.


```{r, results='hide', message=FALSE, warning=FALSE}

tryCatch({
    codebook <- read.csv("~/Sites/gymnast/data/skill-aquirability-variable_key.csv")
    data <- read.csv("~/Sites/gymnast/data/skill-aquirability.csv")
}, error = function(e){
    codebook <- 
        read.csv( github_base_path %+% "data/skill-aquirability-variable_key.csv")
    data <- 
        read.csv( github_base_path %+% "data/skill-aquirability")
})

```

## Display Data
To give a flavor of what the codebook and data look like.

```{r, message=FALSE, warning=FALSE, results='asis'}

data[,1:10] %>% 
    head() %>%
    util.html_table()

codebook %>% 
    head() %>%
    util.html_table()

```

## Using the ds.summarize_by_column function
The summarize by column (`ds.summarize_by_column`) function is used to create
univariate descriptions of data. 

### Default Behavior
By default, `ds.summarize_by_column` uses the functions in
`ds.helper$default_col_funcs` to summarize data.

```{r, message=FALSE, warning=FALSE, results='asis'}

ds.summarize_by_column(data) %>% 
    head() %>%
    util.html_table()

```

### Append a Codebook
If you pass a codebook in the `codebook` argument,
`ds.summarize_by_column` will automatically merge it in on 
the `variable_name` column. Columns from `data`
are retained even if they are not matched in the 
codebook.

```{r, message=FALSE, warning=FALSE, results='asis'}

codebook_columns_to_include <- c("variable_name","variable_label")
ds.summarize_by_column(data, codebook=codebook[, codebook_columns_to_include]) %>% 
    head() %>%
    util.html_table()

```

### Customize Summary Variables
Supply your own summary functions to `ds.summarize_by_column` via the 
`func_list` argument if you want to change the summary.
For example, here we include the mean, standard deviation,
and standard error.

```{r, message=FALSE, warning=FALSE, results='asis'}

custom_func_list <- list(
    "mean" = ds.helper$mean,
    "sd" = ds.helper$sd,
    "se" = function(x){
        # add in the standard error
        x <- x[!util.is_blank(x)]
        sd(x)/length(x)**.5
    }
)

ds.summarize_by_column(data, 
            func_list = custom_func_list,
            codebook=codebook[, codebook_columns_to_include]) %>% 
    head() %>%
    util.html_table()

```

### Use Default Categorical Descriptives
Default description functions for categorical variables are provided in 
`ds.helper$default_categorical_col_funcs` and can be passed to 
the `func_list` argument to summarize categorical data.

```{r, message=FALSE, warning=FALSE, results='asis'}

# select the categorical columns for description
categorical_columns <- ! util.apply_columns(data, util.is_vector_of_numbers)
ds.summarize_by_column( 
        data[ , categorical_columns ], 
        func_list=ds.helper$default_categorical_col_funcs) %>% 
    head() %>%
    util.html_table()

```

## Build and Summarize Simple GLM Models (glm1s)
The `ds` module creates and summarizes glm models when supplied with
outcomes and predictors. It currently only supports models that follow 
the ds.glm1 spec. Canonical ds.glm1 formulas follow this pattern:

```
dv ~ iv [ * mod ] [ + cov1 + cov2 + ... ]
``` 

The ds.glm1 spec is intended to increase predictability and reduce 
summary complexity by restricting certain specifications.
For example, it does not allow multi-level variables in the iv or mod because 
that would require another level of nesting within each model summary.
Only zero or one interactions can be supplied by formula. 

For more on ds.glm1 spec, see comments in `ds.validate_glm1_model()`.




### Build Formulas with ds.build_glm1_formulas()
`ds.build_glm1_formulas()` builds a vector of model formulas for 
each unique combination of the `dv` x `iv` x `mod` x `cov_group`. 
If `cov_group` is supplied with vector,
cov_group implicitly adds one entry for no covs (unadjusted).

```{r, message=FALSE, warning=FALSE, results='asis'}
dvs = c("sksh_post","bls_link") # dependent variables each run independently
ivs = c("condition","skil_aq_1") # independent variables each run independently
mods = c("","in_job_satis")  # moderators each run independently
cov_groups = list(
    c(),  # unadjusted
    c("age","male","race"), # adjust for demographics
    c("sksh_pre","age","male","race") # adjust for demographics + psychometrics
)  # vectors of covariates to run together

ds.build_glm1_formulas( 
            dvs=dvs, ivs=ivs, mods=mods, cov_groups = cov_groups
        ) %>%
    util.print_pre()


```

### Run and summarize with ds.glm1s() and ds.get_model_summary_df
This section models how to use `ds.glm1s()` and `ds.get_model_summary_df`.


```{r, message=FALSE, warning=FALSE, results='asis'}

dvs = c("sksh_post","bls_link") # dependent variables each run independently
ivs = c("treatment") # independent variables each run independently
mods = c("","job_satis")  # moderators each run independently, default no mod
cov_groups = list(
    "unadjusted" = c(),
    "demographic_adjustment" = c("age","male","race")
)  # vectors of covariates to run together
# covs = c()         # covariates included all together or none
                     # here, we use cov_groups instead for more flexibility

data$treatment <- data$condition %in% "treatment"

formulas <- ds.build_glm1_formulas( 
                dvs=dvs, ivs=ivs, mods=mods, cov_groups=cov_groups
            )

# run the models
models     <- ds.glm1s(formulas, data)

# summarize the models into an easy-to-filter df
summary_df <- ds.get_model_summary_df(models, summary_func=ds.summarize_glm1)
# ds.summarize_glm1 extracts many features
names(summary_df) %>% util.print_pre()

# display desired information by filtering the summary_df
# e.g., what models have an iv with p<.01?
# @todo: Create table printing functions with good defaults
summary_df[ summary_df$iv_p < .01, c("formula","iv_apa","int_apa")] %>%
    util.html_table()

# graph the coefficients of the iv with confidence intervals
ggplot(summary_df, 
       aes(formula, iv_coef) ) +
    geom_bar( stat="summary", fun.y="mean", position=ug.dodge) +
    geom_errorbar( 
        aes(ymax = iv_coef + 2 * iv_se, ymin=iv_coef - 2 * iv_se), 
        position=ug.dodge,
        width=0.2
        ) +
    theme_classic() +
    theme(
        legend.key=element_rect(color="black",size=.5),
        axis.text=element_text(face="bold"),
        axis.title=element_text(face="bold")
    ) +
    coord_flip() 

```

### Example of off-spec model
Note that `ds.summarize_glm1` only supports models that fit the ds.glm1 spec.
It will issue warnings if invalid models are provided.
In this example, condition is an invalid iv because it has multiple levels.

```{r, message=FALSE, warning=FALSE, results='asis'}

dvs  = c("sksh_post") # dependent variables each run independently
ivs  = c("condition","treatment") # independent variables each run independently
mods = c("","job_satis")  # moderators each run independently, default no mod
cov_groups = list()  # vectors of covariates to run together
# covs = c()         # covariates included all together or none
                     # here, we use cov_groups instead for more flexibility


formulas <- ds.build_glm1_formulas( 
                dvs=dvs, ivs=ivs, mods=mods, cov_groups = cov_groups
            )

models   <- ds.glm1s(formulas, data)
summary_df <- ds.get_model_summary_df(models, summary_func=ds.summarize_glm1)



```


















